Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2022-05-30 VegOzone.tex
Fix all todos

The results are comprehensively displayed in Fig.~\ref{fig:pody_rel}. Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. The different background colors indicate the temperature response function localization. The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8). Circles indicate simulations with localized GS, canopy height, and leaf dimensions, TODO: (whereas squares indicate simulations with MM GS for Central Europe. We chose this to demonstrate the impact of a projected prolonged GS in the future.) With open symbols, the effect on $\mathrm{POD_y}$ with SWP taken into account is displayed.

We identified in general that an increasing cold tolerance represented by a shift in the temperature response function (MM $<$ cold $<$ subarctic) leads to an increase in $\mathrm{POD_1}$ in both probed years. TODO: (while using locally adapted GS causes a reduction.) An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details). Taking into account SWP, $\mathrm{POD_1}$ was effectively only reduced for deciduous trees in 2018, while in all other cases simulation results with and without SWP effects were identical. An earlier opening of stomata at low light intensities (PPFD0.8) caused an increase in $\mathrm{POD_1}$ compared to MM while a later opening (PPFD1.2) reduced $\mathrm{POD_1}$. Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.

For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by TODO: ($5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$) and display the largest spread in $\mathrm{POD_1}$ with respect to $f_\mathrm{light}$. In 2019, the lowest estimated $\mathrm{POD_1}$ (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, while the highest estimate (subarctic, PPFD0.8) reaches about $20\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. In 2018, this uncertainty range is much smaller ($(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the impact of $f_\mathrm{light}$ is pronounced (extent of $\angle(+,0,-)$). The maximum difference in $\mathrm{POD_1}$ due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019. The difference in $\mathrm{POD_1}$ compared to MM in response to the localization of $f_\mathrm{T}$ and GS ranges between $\Delta\mathrm{POD_1}=(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $\Delta\mathrm{POD_1}=(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in cases when using PPFD0.8. This indicates that drought conditions can reduce the ozone damage risk. TODO: (With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019. However, taking the localized GS into account, this is reversed for the subarctic temperature response function. This can be explained by more favorable growing conditions in 2019.)

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Our localization does not show much difference compared to MM indicating the the MM Boreal Norway spruce parameterization was already well-adjusted to the local subarctic climate. We find a higher $\mathrm{POD_1}$ in 2019 than in 2018. The maximum difference in $\mathrm{POD_1}$ determined between localized GS for subarctic and cold parameterizations with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum difference determined from PPFD0.8 (superscript) and PPFD1.2 (subscript) $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with localize GS stay below the CL. Perennial grassland shows the smallest $\mathrm{POD_1}$, but a similarly large response to the temperature parameterization localizations as deciduous trees. Again, we find a reversal for predicted ozone risk in 2018 and 2019 for the subarctic type. Perennial grassland shows the lowest sensitivity to the light threshold. TODO: (and SWP is only relevant without localized GS (see above).) The overall difference in $\mathrm{POD_1}$ determined from the localized GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The chosen lower maximum temperature in the subarctic parameterization probably caused the lower $\mathrm{POD_1}$ in 2018 (warm year) compared to the cold parameterization. This was not seen in 2019. The maximum difference in $\mathrm{POD_1}$ determined from the PPFD0.8 (superscript) and PPFD1.2 (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the difference between the two probed years. The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but is by construction always larger than estimated with the MM parameterization. This means that the ozone damage risk could be possibly underestimated even in normal years and establishing dedicated subarctic parameterizations are a must to properly assess ozone damage risk in these regions.

>>>>

The results are comprehensively displayed in Fig.~\ref{fig:pody_rel}. Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. The different background colors indicate the temperature response function localization. The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8). Circles indicate simulations with localized GS, canopy height, and leaf dimensions. With open symbols, the effect on $\mathrm{POD_y}$ taking SWP (drought) into account is displayed.

We identified in general that an increasing cold tolerance represented by a shift in the temperature response function (MM $<$ cold $<$ subarctic) leads to an increase in $\mathrm{POD_1}$ in both probed years compared to the MM parameterization. The more extreme the temperature response function localization the higher $\mathrm{POD_1}$. In 2018, however, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to cold in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures. An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details on the particular features of these two years). Taking into account SWP, $\mathrm{POD_1}$ was only notably reduced in 2018 (for deciduous trees). In all other model experiments SWP was hardly of relevance. An earlier opening of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a later opening (PPFD1.2) reduces $\mathrm{POD_1}$. Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.

For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by ($5-16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. In 2018, this range is smaller ($(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). The sensitivity to $f_\mathrm{light}$ is largest in the subarctic parameterization and amounts to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. Herein, the difference of PPFD0.8 compared to MM $f_\mathrm{light}$ is denoted as superscript and for PPFD1.2 as subscript. The extent of the opening angle ($\angle(+,0,-)$) indicates the sensitivity visually. A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in particular when using PPFD0.8. This indicates that drought conditions can reduce the ozone damage risk.

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. The different localizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate. We find higher $\mathrm{POD_1}$ in 2019 than in 2018. The maximum difference in $\mathrm{POD_1}$ determined between subarctic and cold parameterizations with respect to MM amounts to $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum difference due to $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations stay below the CL. Perennial grassland shows the smallest $\mathrm{POD_1}$, but a similarly large response to the temperature parameterization localizations as deciduous trees. Perennial grassland shows the lowest sensitivity to the light threshold. SWP is hardly relevant. The difference in $\mathrm{POD_1}$ for subarctic and cold types with respect to MM amounts to $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum difference in $\mathrm{POD_1}$ determined from the PPFD0.8 (superscript) and PPFD1.2 (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ (2018) and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ (2019), respectively.

The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years and is of the same order of magnitude. By construction, $\mathrm{POD_1}$ from localized parmeterizations is always larger than estimates with the MM parameterization. This means that the ozone damage risk could be possibly underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details). Establishing dedicated subarctic parameterizations are a must to properly assess ozone damage risk in these regions.
